\Title

### 2021-05-31 latex/paper/main.tex
First draft of the proof of lemma, which is in the proof of the theorem

% TODO: remove onecolumn and draftcls formatting
\documentclass[journal,twoside,onecolumn,draftclsnofoot,web]{ieeecolor}

@@

%%
\textbf{Notation}:
%%
We use the following notation and terminology in this article.
The \(n \times n\) identity matrix is denoted \(\Id{n}\), and the \(n \times m\)
matrix of zeros is denoted \(\Zmat{n\times m}\).
A matrix \(A \in \R^{n \times m}\) is \textit{right semi-orthogonal} if
\(A A\tpose = \Id{n}\) and is \textit{left semi-orthogonal} if 
\(A\tpose A = \Id{m}\).
For \(A \in \R^{n\times m}\) and \(B \in \R^{p \times m}\),
we define \([A;B] \in \R^{(n+p)\times m}\) as the matrix obtained by stacking \(A\)
on top of \(B\). 
Given \(\sigma_1,\ldots,\sigma_n \in \R\), we define 
\(\diag{\sigma_1,\ldots,\sigma_n} \in \R^{n \times n}\) as the diagonal matrix
whose value at row \(i\), column \(i\) is \(\sigma_i\).
For \(T > 0\), the set of real numbers modulo \(T\) is denoted \(\Rt{T}\), with
\(\Rt{\infty} := \R\).
The gradient of a matrix-valued function 
\(A : \R^m \rightarrow \R^{n\times n}\) is the block matrix of stacked partial
derivatives, 
\(\nabla_xA := [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
Kronecker product \(A \otimes B \in \R^{nr \times ms}\) (see \cite{kronprod}) is
the matrix
\begin{equation}\label{eqn:kronprod}
    A \otimes B = \begin{bmatrix}
        A_{1,1}B & \cdots & A_{1,m} B \\
        \vdots & \ddots & \vdots \\
        A_{n,1} B & \cdots & A_{n,m} B
    \end{bmatrix} 
    .
\end{equation}
The Poisson bracket \cite{landau_mechanics} between the functions
\(f(q,p)\) and \(g(q,p)\) is
\begin{equation}\label{eqn:poisson-bracket}
    [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
        \pdiff{f}{q_i}\pdiff{g}{p_i}
    .
\end{equation}
Finally, the Kronecker delta \(\delta_i^j = 1\) if \(i = j\) and \(0\)
otherwise.

@@

\begin{lemma}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(\underline{r}, \overline{r} \in \R_{> 0}\) with 
    \(\underline{r} < \overline{r}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to T(D) = ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} = 0\) and \(\dot{\theta} > 0\) on \(T(D)\) when 
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) = 0\).
    Define \(b(\theta,r_0) = g(\theta,r_0,0)\) and assume there exists
    \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(0 < I \leq I^\star\), and lose energy on \(D\)
    for all \(-I^\star \leq I < 0\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0,0) 
    + R(\theta,r_0,I)\) where

    \textbf{TODO: finish the proof}
\end{proof}
\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
that orbits go through Epi}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}

>>>>

% TODO: remove onecolumn and draftcls formatting
\documentclass[journal,twoside,web]{ieeecolor}

@@

%%
\textbf{Notation}:
%%
We use the following notation and terminology in this article.
The \(n \times n\) identity matrix is denoted \(\Id{n}\), and the \(n \times m\)
matrix of zeros is denoted \(\Zmat{n\times m}\).
A matrix \(A \in \R^{n \times m}\) is \textit{right semi-orthogonal} if
\(A A\tpose = \Id{n}\) and is \textit{left semi-orthogonal} if 
\(A\tpose A = \Id{m}\).
For \(A \in \R^{n\times m}\) and \(B \in \R^{p \times m}\),
we define \([A;B] \in \R^{(n+p)\times m}\) as the matrix obtained by stacking \(A\)
on top of \(B\). 
Given \(\sigma_1,\ldots,\sigma_n \in \R\), we define 
\(\diag{\sigma_1,\ldots,\sigma_n} \in \R^{n \times n}\) as the diagonal matrix
whose value at row \(i\), column \(i\) is \(\sigma_i\).
For \(T > 0\), the set of real numbers modulo \(T\) is denoted \(\Rt{T}\), with
\(\Rt{\infty} := \R\).
The gradient of a matrix-valued function 
\(A : \R^m \rightarrow \R^{n\times n}\) is the block matrix of stacked partial
derivatives, 
\(\nabla_xA := [\pdiff{A}{x_1};\ldots;\pdiff{A}{x_m}] \in \R^{nm \times n}\).
Given two matrices \(A \in \R^{n \times m}\) and \(B \in \R^{r \times s}\), the
Kronecker product \(A \otimes B \in \R^{nr \times ms}\) (see \cite{kronprod}) is
the matrix
\begin{equation}\label{eqn:kronprod}
    A \otimes B = \begin{bmatrix}
        A_{1,1}B & \cdots & A_{1,m} B \\
        \vdots & \ddots & \vdots \\
        A_{n,1} B & \cdots & A_{n,m} B
    \end{bmatrix} 
    .
\end{equation}
The Poisson bracket \cite{landau_mechanics} between the functions
\(f(q,p)\) and \(g(q,p)\) is
\begin{equation}\label{eqn:poisson-bracket}
    [f,g] := \sum \limits_{i=1}^n \pdiff{f}{p_i}\pdiff{g}{q_i} - 
        \pdiff{f}{q_i}\pdiff{g}{p_i}
    .
\end{equation}
The Kronecker delta is \(\delta_i^j = 1\) if \(i = j\) and \(0\)
otherwise.
Finally, we say a function \(f(x,I)\) is \(O(I^2)\)
if
\[
    \lim\limits_{\epsilon \to 0} \frac{f(x,I)}{I} = 0
    .
\]

@@

\begin{lemma}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    parameterized by \((q_u,p_u) \in \SxR\).
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    satisfying \(\dot{r} \equiv 0\) and \(\dot{\theta} > 0\) when
    \(I = 0\).
    Let \(g(\theta,r,I) = \dot{r}/\dot{\theta}\) and assume that 
    \((\partial g/\partial I)(\theta,r,0) \equiv 0\).
    Define \(b(\theta,r_0) = (\partial g/\partial r)(\theta,r_0,0)\);
    assume it is \(2\pi\)-periodic in \(\theta\) for fixed \(r_0\),
    and that there exists \(\epsilon > 0\) such that 
    \[
        \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
        ,
    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, ]0,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE 
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \frac{\dot{r}}{\dot{\theta}} =
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    where the prime denotes differentiation by \(\theta\) and 
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) for all \(\theta\),
    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
    \begin{equation}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        \frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder \(R(\theta,r_0,I)\) is \(O(I^2)\).

    By assumption, \((\partial g/\partial I)(\theta,r_0,0) \equiv 0\) and
    \(b(\theta,r_0) := (\partial g/\partial \hat{r})(\theta,r_0,0)\). 
    Hence, one can compute
    \[
        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
        .
    \]
    Since and \(\dot{\theta} > 0\) and \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\);
    this map is 
    \begin{align*}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        .
    \end{align*}
    This map expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\), which
    means we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
    Here, \(r_n\) corresponds to the distance along the line segment
    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
    when an orbit of the constrained dynamics hits \(L\) for the
    the \(n^\text{th}\) time, assuming the orbit was initialized at
    \((r,\theta) = (r_0,0)\).

    Proving the constrained dynamics gain energy on \(D\) is equivalent to
    proving \(\hat{r}_{n}\) eventually reaches \(\overline{r}\).
    Thanks to a perturbation analysis from \cite[Theorem 10.1]{khalil_nonlinear}, 
    we know there exists some \(I_2 \in ]0,I_1]\) and \(\delta > 0\)
    such that \(\left|R(2\pi,r_n,I)\right| \leq \delta I^2\) for all 
    \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0)\) is uniformly
    bounded below by \(\epsilon\), we can find a lower bound on
    \eqref{eqn:r-discrete}:
    \[
        r_{n+1} \geq r_n + I\epsilon - \delta I_2^2
        .
    \]
    Choosing \(I^\star \in ]0,I_2]\) small enough makes 
    \(I\epsilon - \delta I_2^2 \geq \gamma > 0\) for all \(I \in ]0,I^\star]\).
    Therefore, \(r_{n+1} \geq r_n + \gamma \geq r_0 + n\gamma\), which means all
    solutions of \eqref{eqn:r-discrete} flow towards \(\overline{r}\).
    We conclude that, for \(I \in ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).
    Changing the sign of \(I\) means \(r_{n+1} \leq r_n - n\gamma\), in which
    case the constrained dynamics lose energy on \(D\).
\end{proof}
\textbf{TODO: Prove origin is unstable, then apply lemma to O1 and O2 and show
that orbits go through Epi}
%/========== Conclusion ==========/%
\section{Conclusion}\label{sec:conclusion}
